Computational hysteresis:\\ inter-task effects in human computation

### 2014-11-17 docs/publications/2014.09_SCIENCE/new_man.tex
work on manuscript

\paragraph{Effects on workers' direction of focus} 
\begin{figure}
	\centering
	\includegraphics[scale=1]{figs/delta_food.pdf}
	\caption{
		Difference in priming observed for various pairs of worker populations.
		The pairs of populations differed in terms of the initial tasks they
		performed (bars labeled by \textit{inter-t.}) or in terms of the framing
		to which they were exposed (bars labeled by \textit{frame}; see table 1 for
		details about the treatments).  The difference in priming was 
		measured based on the accuracy of a naive Bayes classifier in 
		resolving which treatment workers belonged to, as described in the 
		text.  
	}
	\label{fig:food}
\end{figure}

It is reasonable to expect that exposing a worker to tasks containing images
of food would increase her propensity to provide food-related labels in 
subsequent tasks.  To test whether this straightforward elicitation occured,
we used the wordnet corpus to operationalize the notion of food-related 
labels.  While the coverage of most western food-related words is quite good
we found that the names of many ethnic foods are missing, so we augmented
the corpus by crawling the world food section of the recipes website 
allrecipes.com.  We discuss the validation of our operationalization of 
food-words in detail in the supplementary material.

For \texttt{task1} and \texttt{frame1} we did find that exposing workers to 
food-oriented primes led them to increase their provision of food-related 
labels by about 2-3\% (see Fig.~\ref{fig:food}).  
In \texttt{echo} this effect was stronger
\texttt{echo:food} provided about 8\% more food-related words than 
\texttt{echo:obj}.  Indeed, the \%-increase in food labels observed for
\texttt{echo} was significantly larger than that for \texttt{taks1}, which is 
interesting
given that the overall priming effect observed for \texttt{task1} and 
\texttt{echo} were on par (c.f. Fig.~\ref{fig:theta}).
This makes sense, given that \texttt{echo:food} explicitly calls for a focus 
on food, whereas \texttt{task1:food} probably works by activating certain 
memories and concepts, with a more complex impact on labels.

In \texttt{exp2} we see a strikingly different effect: workers initially 
exposed to food-related tasks actually show fewer food-related labels for the
test tasks.  Recalling the content of the test tasks helps to explain this
observation.  The test tasks in \texttt{exp2} embed both the prime concepts
of food and culture, which we accomplished by using images of food having
very particular and identifiable cultural origins.  The initial tasks in
\texttt{task2:food} were, by comparison, culturally ambiguous, and the worker
would presumably be struck by the appearance of strong cultural overtones
in the test tasks.  On the other hand, due to the effect of negative priming,
workers in \texttt{task2:cult} might be somewhat desensitized to the cultural
content in the images, and instead be struck by the sudden appearance of food.

>>>>

\paragraph{Effects on workers' direction of focus}

Having observed significant effects from framing and task-exposure,
we looked in greater detail at the nature of the effects.  For instance,
does a food-oriented exposure induce workers to provide more food-related
labels?  And if so, does this explain the classifier's ability to distinguish
a worker's initial exposure?

To answer these questions we used the wordnet corpus, augmented with names of
ethnic foods learned from the Internet.  Wordnet provides a (roughly)
hierarchical set of relationships between nouns, called hypernym-hyponym
relationships.  A hypernym is a generalization, while a hyponym is a 
specialization.  So, for example, ``bread'' is hypernym to ``rye bread'' 
and hyponym to ``baked goods''.

Using the hypernym-hyponyms structure of wordnet to identify nouns depicting
food, we found that, for the experimental pairings in \texttt{exp1}, the
food-exposed treatments did produce a significantly larger fraction of 
food-related labels (see Fig.~\ref{fig:specificity}A).  While the shift in 
focus with regard to food does help explain how workers differ depending on 
their exposure, it still leaves most of this difference unexplained.
Furthermore, in \texttt{exp2:task2} we see a significant but reversed effect
(see Fig.~\ref{fig:specificity}A).
Given the difference between the food-culture and food-object axes, it should
not be surprising that the effect of exposure is different.  This tells us that
the differences induced by exposure to frames or initial tasks are significant,
but would appear to be the result of countervailing influences, with the
overall outcome dependant on the particular axis of the exposures, we will
come back to this momentarily.

To explore these effects further, we looked more closely at which particular
labels are most suppressed (activated) by a workers exposure.  We found that, 
for all treatment pairs, ``food'' was always the most suppressed label among 
the food-exposed workers.  This is surprising, given that we have just seen 
that the bulk incidence of food-related labels \textit{increased} among 
food-exposed workers in most treatment pairs.
Since ``food'' is the most generic food-related word in the wordnet-based
ontology, it follows that food-exposed workers have opted to provide more
specialized references to food rather than using ``food''.

Considerable thought must be given to shaping the the vocabulary with which
a worker engages a task.  To draw an example from the domain of citizen 
science, in galaxy zoo, workers catalogue images of galaxies in terms of
smoothness, rings, dustlanes, digital disturbances, and so on.  Shaping this
vocabulary, is accomplished through a combination of positive and negative
examples during training, peer-to-peer and peer-to-expert engagement in 
discussion areas, and using the interface to constrain and prompt the 
worker.

In our set-up, where the worker's vocabulary is unconstrained, we can 
directly see the effects that frames and initial tasks induce on worker 
vocabulary.
By looking at the number of unique food-related labels that workers 
produced, we assesed the richness of vocabulary that workers used when 
referring to food (see Fig.~\ref{fig:specificity}B.  We found that the 
food-exposed workers use a richer vocabulary of 
food-related words, even in cases like \texttt{task2:food}, where the 
food-exposed treatment made fewer references to food overall.  
Interestingly, in this respect, initial taske exposure had a stronger effect
than framing, including the \texttt{echo} treatment pair. This makes sense, 
given that the various images shown in the initial
tasks would tend to activate a variety of memories and concepts related to 
food, reducing the threshold for later activation during the test-tasks.

Using the ontology of food induced by wordnet, we perform a more detailed
calculation of the relative specificity of workers from different treatments.
To do this, we considered all the labels from one treatment
(for example \texttt{task1:food}) and compared these to all the labels from 
another (\texttt{task1:obj}).  For all possible comparisons, we observed 
which treatment's label was more specific more often.  We normalied this
as a percentage (details in supplementary material) so that we can compare the
relative change in specificity between different treatment pairs  .

Performing the above-mentioned calculation while restricting focus to 
food-related labels (see Fig.~\ref{fig:specificity}C), we found that 
the food-exposed workers always provided food-references of substantially 
greater specificity.  In this respect, the effect induced by 
echoed framing was stronger than that induced by the initial task exposures in
\texttt{task1}.

Taking the all the results shown in Fig.~\ref{fig:specificity} together,
a picture emerges which explains the differing natures of intertask and framing
effects.  First, both negative and positive priming would appear to be at play.
Negative priming occurs when a repeated stimulus, which is perceived to be
nonsalient, begins to be ignored.  As a worker from \texttt{task1:food}
proceeds through the initial tasks, food-related memories and concepts are
activated, reducing increasing the likelihood that she provides a variety of
different food-related labels.  This increases the number of food-related
labels, and increases the richness and specificity of her food-related 
vocabulary.  But meanwhile, in the most general sense, the fact that the 
pictures contain food becomes negatively primed.  She does not regard ``food''
to be a meaningful label, and is inhibited in providing it.   
As for the worker in \texttt{task2}, similar effects occur as she proceeds
through the initial tasks.  Upon viewing the first test task, the introduction
of a very prominent cultural element is novel, and is attended to, whereas
her counterpart in \texttt{task2:cult} has seen several images containing 
cultural themes, and is negatively primed to the general setting of culture,
but responds to the novelty of the introduction of food.  Finally, looking at
the results for \texttt{echo}, we see evidence of its effects operating through
the executive faculties of the worker.  There is considerably more
restriction to food-related labels, and even a greater specificity of such 
labels, but, lacking the diverse activations of food-related memories and 
concepts from initial tasks, the richness of vocabulary is lower.

Based on these observations, we submit several recommendations for the 
consideration for the designers of croudsourced initiatives.
\begin{enumerate}
	\item{
		Examples provide a valuable way to clarify intstructions, but 
		training tasks should be preferred, since they activate concepts and 
		faculties relevant to the subsequent task.
	}
	\item{
		Examples used for training should not be too ``clean'', but should 
		contain the same distribution of noise, or distracting features, as 
		is observed in the real dataset.
		This may help desensitize the worker to the nonsalient features.
		Using training tasks that are too clean will make the subsequent noise
		more novel and thus more distracting.
	}
	\item{
		It is common practice to plant \textit{validation tasks}, for which the
		correct answer is known, to verify worker reliability.
		These tasks should simultaneously be considered as 
		\textit{callibration tasks}, which help
		the worker to maintain an optimally primed state.
		Giving feedback, for both correct and incorrect callibration tasks, 
		may help further reinforce the salient features for the tasks.
	}
	\item{
		Some tasks require the classification of imbalanced datasets.
		For example, in the detection of pre-ictal (pre-siezure) EEG traces,
		the majority of traces are normal.  The absence of exposure to the 
		positive class can lead to a loss of activation for the
		salient features and a drifting conceptualization of the class.
		Using callibration tasks drawn from the underrepresented class may 
		help sustain workers' attention and recall for that class.
	}
	\item{
		When a very precice classification is needed, it may be helpful to
		performing the classification in multiple steps.  For example, in a 
		binary classification, the workers could initially perform a coarse
		classification, using three options for response: 
		``definitely class 0'', ``definitely class 1'', and 
		``borderline''.  Following the coarse classification, the 
		``borderline'' examples could be served in batches together, 
		allowing workers to focus on finer distinctions which are salient 
		in the context of borderline examples.
	}
\end{enumerate}

Each of these techniques can and should be tested directly, and the the
lessons derived from those experiments should be encorporated into the best 
practices for human computation.

\paragraph{Conclusion.}
Our results definitively show that intertask effects can
be as strong as framing effects, and appear in general to be stronger.
Intertask effects have are similar in strength to direct 
statements about the purpose of the tasks, followed by the reiteration of the
purpose by the worker.  However, they differ in the detailed effects that
they bring about.  Initial tasks had a greater effect on richness of 
spontaneously elicited vocabulary used to refer to the primed concept,
whereas the echoed purpose had a somewhat stronger effect increasing the 
quantity of references to the primed concept.

The task designer would do well to give consideration to intetask effects
when performing task batching, and when designing training and validation
tasks.  Despite all efforts to eliminate biases in the microtask setup,
the greatest source of biases may lurk in the tasks themselves.  
But the task designer need not seek to factor out intertask effects altogether.
Rather we advocate regarding intertask effects as a tunable parameter which
can be used to optimize HPU performance.  By correctly directing attention
and conceptual activation, intertask effects may be helpful in exceeding
expert annotations using non-expert HPU computing in a wider variety of
applications.
